---
title: "Hands-on Exercise 8B: Visualising Geospatial Point Data"
author: "Kristine Joy Paas"
date: "7 June 2024"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

This hands-on exercise covers [Chapter 22: Visualising Geospatial Point Data](https://r4va.netlify.app/chap22).

In this exercise, I learned:

-   To import an aspatial data file into R.

-   To convert it into simple point feature data frame and at the same time, to assign an appropriate projection reference to the newly create simple point feature data frame.

-   To plot interactive proportional symbol maps.

# 2 Getting Started

## 2.1 Loading the required packages

For this exercise we will use the following R packages:

-   **sf**: for processing geospatial data

-   **tidyverse**: data analytics tools for r

-   **tmap**: for generating maps

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## 2.2 Importing data

The data set use for this hands-on exercise is called *SGPools_svy21*. The data is in csv file format.

```{r}
sgpools <- read_csv("data/aspatial/SGPools_svy21.csv")
head(sgpools)
```

It contains a list SG Pools outlets, and their corresponding geospatial data (**XCOORD**, **YCOORD**) that we can use to plot them in a map.

# 3 Data Wrangling

To work with `sf` we need to convert this data to the appropriate type:

```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
```

-   `coords`: specify which columns the coordinate information are in

-   `crs`: to provide the coordinates system in epsg format. [EPSG: 3414](https://epsg.io/3414) is Singapore SVY21 Projected Coordinate System. You can search for other country’s epsg code by refering to [epsg.io](https://epsg.io/).

```{r}
head(sgpools_sf)
```

This contains a `geometry` column which contains the geospatial representation of the provided coordinates. In this case. there are `point` data.

# 4 Drawing Proportional Symbol Map

To create an interactive proportional symbol map in R, the view mode of tmap will be used.

The code churn below will turn on the interactive mode of tmap.

```{r}
tmap_mode("view")
```

## 4.1 Plotting an interactive point symbol map

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = 1,
           border.col = "black",
           border.lwd = 1)
```

## 4.2 Proportional point sizes

The bubbles in the map above have the same sizes. Let's make the bubbles bigger the higher the winnings are.

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = "Gp1Gp2 Winnings",
           border.col = "black",
           border.lwd = 1)
```

## 4.3 Assign colors

The proportional symbol map can be further improved by using the colour visual attribute. In the code chunks below, *OUTLET_TYPE* variable is used as the colour attribute variable.

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1)
```

## 4.4 Adding facets

The argument *sync* in *tm_facets()* can be used to produce multiple maps with synchronised zoom and pan settings.

```{r}
tm_shape(sgpools_sf) +
  tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1) +
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

Switch back to non-interactive mode

```{r}
tmap_mode("plot")
```

# 5 Reflections

During ISSS624, I rarely use interactive plots as it was using up a lot for resources. However, doing it the non-interactive way was more tedious as I had to do additional wrangling of the map use in the background (in this case, Singapore map).

However, the interactive map has a lot of features that I like (e.g. synchronized facets) so I'll try to use it more.
