---
title: "In-class Exercise 5: Visualizing and Analyzing Text Data"
author: "Kristine Joy Paas"
date: "11 May 2024"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

We will visualize and analyze text data

# 2 Getting Started

## 2.1 Loading the required packages

-   readtext - Import and handling for plain and formatted text

-   quanteda - Quantitative text data analysis

-   tidytext - facilitates use of text to other tidy formats

-   jsonlite - for processing json files

```{r}
#| label: setup
pacman::p_load(tidyverse, readtext, quanteda, tidytext, jsonlite, DT)
```

## 2.2 Loading the data

Using `readtext` reads all files and add the contents in tibble dataframe.

We use the `articles/` and `mc1.json` data from [VAST Challenge](https://vast-challenge.github.io/2024/MC1.html).

```{r}
text_data = readtext("data/articles/*")
glimpse(text_data)
```

# 3 Tokenizing text data

This will extract the words/token from the text data

```{r}
usenet_words <- text_data %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)
glimpse(usenet_words)
```

## 3.1 Sorting data by frequency

We can sort the words by frequency by using `count`.

```{r}
usenet_words %>%
  count(word, sort = TRUE)
```

The 3 most common words from the text sources are `fishing`, `sustainable`, and `company`.

## 3.2 Extracting company and news source information

The file name seems to be in the format `CompanyName__0__Publisher`. We can extract company name and publisher.

```{r}
text_data_split <- text_data %>%
  separate_wider_delim("doc_id",
                       delim = "__0__",
                       names = c("X", "Y"),
                       too_few = "align_end")
glimpse(text_data_split)
```

::: callout-tip
The code above hardcodes `0` in the delimiter but it can be any number. As a result, some are not parsed as expected, e.g.

|     |                                                   |     |
|-----|---------------------------------------------------|-----|
| NA  | Cervantes-Kramer\_\_1\_\_1\_\_Haacklee Herald.txt | C   |

Looking at the data a second time, we can determine that the delimiter should be something like `__<num>__<num>__`.
:::

We can make adjustments to the code to use `regex` to use regular expression with `\d+` token to match any number. We will also change "X" and "Y" to "Company" and "Publisher", respectively so we can more easily understand what the data represents.

```{r}
text_data_company_publisher <- text_data %>%
  separate_wider_delim("doc_id",
                       delim = regex("__\\d+__\\d+__"),
                       names = c("Company", "Publisher"),
                       too_few = "align_end")
datatable(text_data_company_publisher[1:2])
```

# 4 Working with json files

We can use `fromJSON` to read json file

```{r}
mc1_data <- fromJSON("data/mc1.json")
```

# 5 Reflections
