---
title: "Hands-on Exercise 10: Information Dashboard Design: R methods"
author: "Kristine Joy Paas"
date: "18 June 2024"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

This hands-on exercise covers [Chapter 31: Information Dashboard Design: R methods](https://r4va.netlify.app/chap31).

In this exercise, I learned:

-   Building dashboard

# 2 Getting Started

## 2.1 Loading the required packages

For this exercise we will use the following R packages:

::: {.callout-tip appearance="simple"}
### Packages added

-   `svglite` - when plotting bullet chart with `gt`.
:::

```{r}
pacman::p_load(lubridate, ggthemes, reactable,
reactablefmtr, gt, gtExtras, tidyverse, svglite)
```

## 2.2 Importing data

For the purpose of this study, we will import RDS file `CofeeChain.rds`.

```{r}
coffeechain <- read_rds("data/rds/CoffeeChain.rds")
```

## 2.3 Exporting to CSV

::: callout-important
### .mdb not working on Mac

As `.mdb` does not work on Mac, I exported the data to CSV to make it work with Tableau
:::

```{r}
write.csv(coffeechain,"data/CoffeeChain.csv")
```

# 3 Plotting with ggplot

## 3.1 Bullet chart

We will aggregate Sales and Budgeted Sales at the Product level.

```{r}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

Plot using ggplot

```{r}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

## 3.2 Plotting sparklines

### 3.2.1 Preparing data

Generate the sales report

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

Compute the minimum, maximum and end of month sales

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

Compute 25th and 75th quantiles

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)
```

### 3.2.2 Generating chart

```{r}
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

# 4 Static Information Dashboard Design: gt and gtExtras methods

We will create a static information dashboard using `gt` and `gtExtras`.

## 4.1 Bullet chart

::: {.callout-note appearance="simple"}
### Packages added

Added `svglite` so the plot can be generated.
:::

We will now plot the chart using `gt`.

```{r}
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## 4.2 Sparklines using gtExtras

### 4.2.1 Preparing sales data

```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()

head(report)
```

We need a list type for the sales column so we need to further process the data.

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), .groups = "drop")
```

### 4.2.2 Plotting sparklines

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)
```

### 4.2.3 Calculating statistics

We will generate statistics to add to the plot.

```{r}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

### 4.2.4 Combining sparklines and statistics

```{r}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

```{r}
sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )
```

```{r}
sales_data = left_join(sales, spark)
```

### 4.2.5 Plotting the updated data table

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```

### 4.2.6 Combining bullet chart and sparklines

```{r}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 
```

```{r}
sales_data = sales_data %>%
  left_join(bullet)
```

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

# 5 Interactive Information Dashboard Design: reactable and reactablefmtr methods

We will create interactive information dashboard by using [reactable](https://glin.github.io/reactable/) and [reactablefmtr](https://kcuilla.github.io/reactablefmtr/) packages.

In order to build an interactive sparklines, we need to install dataui R package.

```{r}
remotes::install_github("timelyportfolio/dataui")
library(dataui)
```

## 5.1 Plotting interactive sparklines

Similar to before, we will perform some data aggregation.

```{r}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

Then, we will use `reactable` to plot the sparklines. We will also set the `defaultageSize` so everything will appear in 1 page.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

## 5.2 Adding points and labels

We will mark the minimum and maximum values in each line.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```

## 5.3 Adding reference line

We will add a reference line for the mean

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```

## 5.4 Adding bandline

We can also opt for a bandline instead of the reference line.

The plot below highlights the inner quartile, which is useful to see how spread out the values are.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```

## 5.5 Using sparkbar

We can also opt to use sparkbar instead of sparkline

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```

# 6 Reflections

I had trouble running this exercise at first due to the Microsoft DB requirement. The tools used are too old and may not be compatible with the computers that the students use. For example, I am using Macbook with Silicon chip, hence Microsoft-specific tools have limited compatibility. On top of that, the tools used (e.g. `odbcConnectAccess2007` works on almost 2 decade old machines), which is not compatible with newer computers, or will need a lot of debugging and research to work. This takes too much time to setup.

Eventually, I was able to find a solution that works for my machine: <https://github.com/ethoinformatics/r-database-connections/blob/master/Access-from-R-Mac.md>. However, this may not be applicable to all. Good thing the `rds` file was released so I can proceed.

I really like the alternative tools used for this exercise: `gt` and `reactable`. This is because they make it very easy to create very aesthetic plots. Although I came to like `ggplot`. It can be a lot of pain to generate the plot I imagine due to its complexity. This is due to it catering to various kind of data and plots.

This is why specialist tools like `gt` and `reactable` that do 1 thing are useful if one's purposes match with their functionalities.
