---
title: "Hands-on Exercise 5: Visualizing and Analyzing Text Data"
author: "Kristine Joy Paas"
date: "3 May 2024"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1 Overview

This hands-on exercise covers [Chapter 29: Visualising and Analysing Text Data with R: tidytext methods](https://r4va.netlify.app/chap29).

I learned about the following:

-   understand tidytext framework for processing, analysing and visualising text data,

-   write function for importing multiple files into R,

-   combine multiple files into a single data frame,

-   clean and wrangle text data by using tidyverse approach,

-   visualize words with Word Cloud,

-   compute term frequency–inverse document frequency (TF-IDF) using tidytext method, and

-   visualizing texts and terms relationship.

# 2 Getting Started

## 2.1 Loading the required packages

For this exercise we will use the following R packages:

-   tidytext, tidyverse (mainly readr, purrr, stringr, ggplot2)

-   widyr,

-   wordcloud and ggwordcloud,

-   textplot (required igraph, tidygraph and ggraph, )

-   DT,

-   lubridate and hms

```{r}
#| label: setup
pacman::p_load(tidytext, widyr, wordcloud, DT, ggwordcloud, textplot, lubridate, hms, tidyverse, tidygraph, ggraph, igraph)
```

## 2.2 Importing Multiple Text Files from Multiple Folders

### 2.2.1 Creating a folder list

```{r}
news20 <- "data/20news/"
```

### 2.2.2 Define a function to read all files from a folder into a data frame

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}
```

## 2.3 Importing Multiple Text Files from Multiple Folders

### 2.3.1 Reading in all the messages from the 20news folder

```{r}
raw_text <- tibble(folder = 
                     dir(news20, 
                         full.names = TRUE)) %>%
  mutate(folder_out = map(folder, 
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), 
            id, text)
write_rds(raw_text, "data/rds/news20.rds")
```

# Reflections
